Object Capability Patterns:\\Policies and Specifications

### 2017-08-18 article.tex
some RNode
RN

\begin{itemize}
\item \textbf{Initial Conditions or Endowment:} o has an existing path to o' in \losigma, therefore X* refers to o'
\item \textbf{Parenthood:} o will create o' in some arising \losigma', therefore X* refers to o
\item \textbf{Introduction:} o will only obtain the path to o' in some arising \losigma', therefore X* refers to an object that is not the same object as o (X*\loneq o). 
\end{itemize}

@@

The contraposition result of \textsc{epc1} says that for there to be no eventual paths from o to o' these configurations must hold:
\begin{itemize}\setlength\itemsep{0.5em}
\item if there is some object x that o has a path to in \losigma, then x cannot have an eventual path to o'
\item if there is some object x that possesses the capability of o in \losigma, then x cannot have an eventual path to o'
\end{itemize}
Note that object x can refer to the same object as o.
There is one final critical result from \textsc{epc1}.
Notice how, there is a 'recursive' MayAccess$^{Ind,Eve}$(X*,o') in our condition for MayAccess$^{Ind,Eve}$(o,o').
This allows us to recursively expand the condition to incorporate \textit{all} X* intermediate objects in the path leading to o'. This recursive expansion gives us both MayAccess$^{Dir,Now}$ and MayAccess$^{Ind,Now}$ predicates, where the terminating MayAccess$^{Ind,Eve}$(o',o') always returns true if o' exists. This result allows us to define MayAccess$^{Ind,Eve}$(X*,o') in a configuration of paths completely based in \losigma.\\

@@

To give a concrete example, let us assume a simple system with only three objects, o, x, o'.
\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Path Example}\\
M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \ablock $[$\null
\textcolor{red}{MayAccess$^{Ind,Now}$(o,o) \loand\linebreak
\ablock \ablock \quad MayAccess$^{Dir,Now}$(o,x)\linebreak
}
\ablock \ablock \quad \loor \linebreak
\ablock \ablock MayAccess$^{Dir,Now}$(x,o)$ ]$\linebreak
\ablock \ablock \loand \linebreak
\textcolor{blue}{
\ablock \ablock $\{[$
\textcolor{red}{MayAccess$^{Ind,Now}$(x,x) \loand\linebreak
\ablock \ablock \quad MayAccess$^{Dir,Now}$(x,o')\linebreak}
\ablock \ablock \quad \quad \loor \linebreak
\ablock \ablock \quad MayAccess$^{Dir,Now}$(o',x)$]$\linebreak
\ablock \ablock \quad \loand \linebreak
\ablock \ablock \quad MayAccess$^{Ind,Eve}$(o',o')$\}$
}\linebreak
\ablock \ablock $\}$
\linebreak \\
\hrule
\end{logic}
In the example, we make concrete \textsc{epc2} using X* = x and Y* = o', where the expansion terminates at the three predicates, MayAccess$^{Ind,Now}$(o,o), MayAccess$^{Ind,Now}$(x,x) and MayAccess$^{Ind,Eve}$(o',o'). Let us assume o, x, and o' always exists, such that these predicates would return true. This example then illustrates an eventual path from o to o' can only exist if one of these four conditions in \losigma\ hold:
\begin{itemize}\setlength\itemsep{0.5em}
\item o has the capability of x, and x has the capability of o'
\item o has the capability of x, and o' has the capability of x
\item x has the capability of o, and x has the capability of o'
\item o has the capability of x, and x has the capability of o'
\end{itemize}
Note how, these four configurations are present state configurations at \losigma, but allows us to reason about whether a potential path from o to o' can exist in all arising states \losigma' from \losigma.\\

So far we have developed results for paths from o to o', but what about a direct path from o to o'? Luckily, we only need construct \textsc{eec} from \textsc{epc1}, the only difference being we now require the path from o to X* in \textsc{epc1} to be traversable.

\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Execution Connectivity (EEC)}\\
M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj.$[($MayAccess$^{Dir,Now}$(o,X*) \loor \linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ 
\loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Dir,Now}$(o,X*) \loand \linebreak
\ablock \ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \ablock \ablock \ablock \loneg $[$MayAccess$^{Dir,Eve}$(X*,o')$]\}$\linebreak
\ablock \loimplies\linebreak
	\ablock\loneg MayAccess$^{Dir,Eve}(o,o')$\linebreak
\hrule
\end{logic}
The power of \textsc{eec} is similar to \textsc{epc}\ in that it allows an arbitrary recursive expansion of the MayAccess$^{Dir,Eve}$(X*,o') till the terminating MayAccess$^{Dir,Eve}$(o',o') that is always true.\\

We can always use an automated theorem prover to prove whether a path can arise between two objects in a present configuration system.
We now turn our attention formalizing Ocap state changes rules in the system.\\

@@

\subsubsection{Protection is about local object-path denial and global capability denial}
\begin{logic}
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \lonimplies 
MayAccess$^{Dir,Eve}$(o,o')\linebreak
\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies 
MayAccess$^{Ind,Eve}$(o,o')\linebreak
\\
M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies 
\loneg MayAffect$^{Eve}$(o,o')\linebreak
\\
M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(o,o') \loimplies 
MayAccess$^{Ind,Eve}$(o,o')\linebreak
\\
M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies 
\loneg MayAccess$^{Dir,Eve}$(o,o') \loand\ \loneg MayAffect$^{Eve}$(o,o')\linebreak
\end{logic}
We make the important result that protection of an object's state is path denial to the object, where path is a necessary condition for an object to affect another object. An object o having the capability of o' is \textbf{NOT} a necessary condition for o to modify the state of o'. However, the existence of \textit{some} object X* (that can refer to o, or other objects) having the capability of o', is a necessary condition for o to modify o'. See \textsc{nec1} and \textsc{nec2} below which illustrates this. This is precisely why attenuating objects are so powerful and useful in OCap patterns. We can always deny o the capability of o', but still allow o to modify o' in some protected way. To say it in another way, denial of paths is much stronger, as it implies protection \textit{and} denial of capability, but denial of capability does \textit{not} imply protection and denial of paths.\\

\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Execution Condition 1 (NEC1)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
	\ablock \loexists X*\loin{Obj}. MayAccess$^{Dir,Eve}$(X*,o') 
\linebreak\\
\textit{By contraposition and quantifier equivalence,}\linebreak
M,\losigma\ \loturns\ $[$\loforall X*\loin{Obj}. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$\loimplies\linebreak
	\ablock \loneg MayAffect$^{Eve}$(o,o') \linebreak
\hrule
\end{logic}
\textsc{nec1} says that in order for an object o to modify the state of o', there must exist an object X* that has a direct path to o', and that X* \textit{can} traverse such a path. This is derived trivially from our \textsc{pfa} assumption. Consequently, in our contraposition result, we can deny \textit{all} objects in our system capability of o' to enforce the protection of o' from o. \textsc{nec1} does not yield a very useful result, as this is just another way of implementing protection of o' from o through denying all paths to o' in the system reference graph. To see why denying all objects the capability of o' is equivalent to denying all paths to o', see the contraposition result from \textsc{gpe}.\\

In \textsc{nec1}, if X* refers to the same object as o, then we have a straightforward configuration where object o has the capability of o' and can therefore affect o'. However, very crucially, o having the capability of o' is \textit{not} a sufficient condition---denying o the capability of o' does \textit{not} deny o the ability to affect o', since X* can refer to an object that is \textit{not} o. \textsc{nec2} below is an expansion of \textsc{nec1} to illustrate this crucial point.
\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Execution Condition 2 (NEC2)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
	\ablock $[$MayAccess$^{Dir,Eve}$(o,o') \loor\linebreak
	\ablock \loexists X*\loin{Obj}, X*\loneq o. MayAccess$^{Dir,Eve}$(X*,o')$]$
\linebreak\\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ $[$\loneg MayAccess$^{Dir,Eve}$(o,o') \loand\linebreak
	\ablock \loforall X*\loin{Obj}, X*\loneq o. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$ \linebreak
	\ablock \loimplies\loneg MayAffect$^{Eve}$(o,o')\linebreak
\hrule
\end{logic}
With this clarification, we then reason about the relationship between protection of state (MayAffect) and paths (MayAccess$^{Ind,*}$). The question we ask is what is the relation between state protection in eventual outcomes, and present path configurations? We build this relationship from the basics by progressively finding stronger conditions of MayAffect. We begin with the first condition, which says that for an object to affect another, an \textit{eventual} path has to exist. This can be derived from the definitions of our predicates.
\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Path Condition 1 (NPC1)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
	\ablock MayAccess$^{Ind,Eve}$(o,o')
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies\linebreak
	\ablock \loneg MayAffect$^{Eve}$(o,o')
\linebreak
\hrule
\end{logic}
Here, we immediately see a first defensive outcome of the OCap model. Having no eventual paths from object o to o' guarantees that the state of object o' cannot be modified by object o.\\

So far our results serve as a good base to enforce \loneg MayAffect$^{Eve}(o,o')$ but we require \textit{stronger} necessary conditions for MayAffect$^{Eve}(o,o')$, because we need to understand the present path configurations that can deny o from modifying o'. Luckily, we have already built the necessary ingredients in our formalizations of OCap rules in \ref{ocaprules}, where have painstakingly derive the relationship between present path-capability configurations and eventual paths. Since \textsc{npc1} tells us that the eventual paths is a necessary condition for object state modification, and \textsc{epc1} tells us the the relationship between eventual paths and present path-capability configurations, we combine the two to present:

\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Path Condition 2 (NPC2)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')$]$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ 
\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
\ablock \loimplies \linebreak
	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
\hrule
\end{logic}
, where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use the condition to expand the condition till we reach the terminating o'. \textsc{npc2} gives us the present path configurations that will allow an eventual path from o to o' as a necessary condition for o to modify o'.But if objects are path-isolated, how do we allow them to cooperate?

\textsc{nec2} and \textsc{npc2} show the necessary capability and path conditions for protection. Is there a relation that can capture the essence of both? Notice how, we have said in \textsc{nec2} that there must exist some object in the system (that is not necssarily o) that has the capability of o', while in \textsc{npc2}, we said that o must have a path to o'. Surely it must mean then that the object X* in \textsc{nec2} that has the capability of o' refers to the last object on the path from o to o' mentioned in \textsc{npc2}. We now bring these results together, by making use of \textsc{gpec} to make explicit the last object in the eventual path holding the capability of o' in \textsc{npc2}:\\
\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Execution-Path Condition* (NEPC*)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \loexists Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loand\linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ 
\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \quad\loforall Y*\loin{Obj}. $[$MayAccess$^{Ind,Eve}$(X*, Y*) \loor\linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$
\ablock \loimplies \linebreak
	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
\hrule
\end{logic}
\textsc{nepc*} is nice, because we can see the necessary path condition and the condition that an object must hold the capability of o'. But what does it mean to modify an object? No matter how long the path from o to o' is, surely to modify o' requires the last object on our path to \textit{be} the 'final messenger' that will call the object's behaviour to modify it---which means MayAffect$^{Eve}$(Y*,o') \textit{must} hold. We express this by extending \textsc{nepc1*} to \textsc{nepc2*}.
\begin{minipage}{\linewidth}
\begin{logic}

>>>>

\begin{itemize}
\item \textbf{Initial Conditions or Endowment:} o has an existing path to o' in \losigma, therefore X* refers to o'
\item \textbf{Parenthood:} since o can create o' in some arising \losigma', then o can also create o' in \losigma, X* refers to o'
\item \textbf{Introduction:} o will only obtain the path to o' in some arising \losigma', therefore X* refers to an object that is not the same object as o (X*\loneq o). 
\end{itemize}

@@

The contraposition result of \textsc{epc1} says that for there to be no eventual paths from o to o' these configurations must hold:
\begin{itemize}\setlength\itemsep{0.5em}
\item if there is some object x that o has a path to in \losigma, then x cannot have an eventual path to o'
\item if there is some object x that possesses the capability of o in \losigma, then x cannot have an eventual path to o'
\end{itemize}
Note that object x can refer to the same object as o'.
There is one final critical result from \textsc{epc1}.
Notice how, there is a 'recursive' MayAccess$^{Ind,Eve}$(X*,o') in our condition for MayAccess$^{Ind,Eve}$(o,o').
This allows us to recursively expand the condition to incorporate \textit{all} X* intermediate objects in the path leading to o'. This recursive expansion gives us both MayAccess$^{Dir,Now}$ and MayAccess$^{Ind,Now}$ predicates, where the terminating MayAccess$^{Ind,Eve}$(o',o') always returns true if o' exists. This result allows us to define MayAccess$^{Ind,Eve}$(X*,o') in a configuration of paths completely based in \losigma.\\

@@

To give a concrete example, let us assume a simple system with only three objects, o, x, o'.
\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Path Example}\\
M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \ablock $[$\null
\textcolor{red}{MayAccess$^{Ind,Now}$(o,o) \loand\linebreak
\ablock \ablock \quad MayAccess$^{Dir,Now}$(o,x)\linebreak
}
\ablock \ablock \quad \loor \linebreak
\ablock \ablock MayAccess$^{Dir,Now}$(x,o)$ ]$\linebreak
\ablock \ablock \loand \linebreak
\textcolor{blue}{
\ablock \ablock $\{[$
\textcolor{red}{MayAccess$^{Ind,Now}$(x,x) \loand\linebreak
\ablock \ablock \quad MayAccess$^{Dir,Now}$(x,o')\linebreak}
\ablock \ablock \quad \quad \loor \linebreak
\ablock \ablock \quad MayAccess$^{Dir,Now}$(o',x)$]$\linebreak
\ablock \ablock \quad \loand \linebreak
\ablock \ablock \quad MayAccess$^{Ind,Eve}$(o',o')$\}$
}\linebreak
\ablock \ablock $\}$
\linebreak \\
\hrule
\end{logic}
In the example, we make concrete \textsc{epc2} using X* = x and Y* = o', where the expansion terminates at the three predicates, MayAccess$^{Ind,Now}$(o,o), MayAccess$^{Ind,Now}$(x,x) and MayAccess$^{Ind,Eve}$(o',o'). Let us assume o, x, and o' always exists, such that these predicates would return true. This example then illustrates an eventual path from o to o' can only exist if one of these four conditions in \losigma\ hold:
\begin{itemize}\setlength\itemsep{0.5em}
\item o has the capability of x, and x has the capability of o'
\item o has the capability of x, and o' has the capability of x
\item x has the capability of o, and x has the capability of o'
\item o has the capability of x, and x has the capability of o'
\end{itemize}
Note how, these four configurations are direct capability configurations at \losigma, but allows us to reason about whether a potential path from o to o' can exist in some arising state \losigma' from \losigma. This is expressed more succintly below in \textsc{epc3***}

\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Path Connectivity 3*** (EPC3***)}\\
M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*, x$^{1...n}$\loin Obj.$\{[$MayAccess$^{Ind,Now}$(o,o)\linebreak
\ablock \ablock \ablock \ablock \ablock \loand\linebreak
\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(o,x$^1$)\linebreak
\ablock \ablock \ablock \ablock \ablock $\vdots$\linebreak
\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(x$^{n-1}$,x$^n$)\linebreak
\ablock \ablock \ablock \ablock \ablock \loand\linebreak
\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(x$^n$,X*)$]$\linebreak
\ablock \ablock \ablock \ablock \quad\: \loor \linebreak
\ablock \ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$\}$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')\linebreak\\
\ablock \textit{where if o exists in \losigma,} \linebreak
\ablock \loturns\ MayAccess$^{Ind,Now}$(o,o) is true\linebreak\linebreak\
\ablock \textit{where if o' exists in some \losigma',}\linebreak
\ablock \loturns\ MayAccess$^{Ind,Eve}$(o',o') is true\linebreak
\hrule
\hrule
\end{logic}

So far we have developed results for paths from o to o', but what about a direct path from o to o' (what are the conditions such that o will have the direct capability of o' eventually)? Luckily, we only need construct \textsc{eec} from \textsc{epc1}, the only difference being we now require the path from o to X* in \textsc{epc1} to be traversable.

\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Execution Connectivity (EEC)}\\
M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj.$[($MayAccess$^{Dir,Now}$(o,X*) \loor \linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ 
\loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Dir,Now}$(o,X*) \loand \linebreak
\ablock \ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \ablock \ablock \ablock \loneg $[$MayAccess$^{Dir,Eve}$(X*,o')$]\}$\linebreak
\ablock \loimplies\linebreak
	\ablock\loneg MayAccess$^{Dir,Eve}(o,o')$\linebreak
\hrule
\end{logic}
The power of \textsc{eec} is similar to \textsc{epc1}\ in that it allows an arbitrary recursive expansion of the MayAccess$^{Dir,Eve}$(X*,o') till the terminating MayAccess$^{Dir,Eve}$(o',o').\\

We now turn our attention formalizing Ocap state changes rules in the system.\\

@@

\subsubsection{Protection is path-denial}
\begin{logic}
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \lonimplies 
MayAccess$^{Dir,Eve}$(o,o')\linebreak
\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies 
MayAccess$^{Ind,Eve}$(o,o')\linebreak
\\
M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies 
\loneg MayAffect$^{Eve}$(o,o')\linebreak
\\
M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(o,o') \loimplies 
MayAccess$^{Ind,Eve}$(o,o')\linebreak
\\
M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies 
\loneg MayAccess$^{Dir,Eve}$(o,o') \loand\ \loneg MayAffect$^{Eve}$(o,o')\linebreak
\end{logic}
We make the important result that protection of an object's state is path denial to the object, where path is a necessary condition for an object to affect another object. An object o having the capability of o' is \textbf{NOT} a necessary condition for o to modify the state of o'. However, the existence of \textit{some} object X* (that can refer to o, or other objects) having the capability of o', is a necessary condition for o to modify o'. See \textsc{ncc1} and \textsc{ncc2} below which illustrates this. This is precisely why attenuating objects are so powerful and useful in OCap patterns. We can always deny o the capability of o', but still allow o to modify o' in some protected way. To say it in another way, denial of paths is much stronger, as it implies protection \textit{and} denial of capability, but denial of capability does \textit{not} imply protection and denial of paths.\\

\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Capability Condition 1 (NCC1)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
	\ablock \loexists X*\loin{Obj}. MayAccess$^{Dir,Eve}$(X*,o') 
\linebreak\\
\textit{By contraposition and quantifier equivalence,}\linebreak
M,\losigma\ \loturns\ \loforall X*\loin{Obj}. \loneg MayAccess$^{Dir,Eve}$(X*,o')\loimplies\linebreak
	\ablock \loneg MayAffect$^{Eve}$(o,o') \linebreak
\hrule
\end{logic}
\textsc{ncc1} says that in order for an object o to modify the state of o', there must exist an object X* that has a direct path to o', and that X* \textit{can} traverse such a path. This is derived trivially from our \textsc{pfa} assumption. Consequently, in our contraposition result, we can deny \textit{all} objects in our system capability of o' to enforce the protection of o' from o. \textsc{nec1} does not yield a very useful result, as this is just another way of implementing protection of o' from o through denying all paths to o' in the system reference graph. To see why denying all objects the capability of o' is equivalent to denying all paths to o', see the contraposition result from \textsc{gpe}.\\

In \textsc{ncc1}, if X* refers to the same object as o, then we have a straightforward configuration where object o has the capability of o' and can therefore affect o'. However, very crucially, o having the capability of o' is \textit{not} a sufficient condition---denying o the capability of o' does \textit{not} deny o the ability to affect o', since X* can refer to an object that is \textit{not} o. \textsc{ncc2} below is an expansion of \textsc{ncc1} to illustrate this crucial point.
\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Capability Condition 2 (NCC2)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
	\ablock $[$MayAccess$^{Dir,Eve}$(o,o') \loor\linebreak
	\ablock \loexists X*\loin{Obj}, X*\loneq o. MayAccess$^{Dir,Eve}$(X*,o')$]$
\linebreak\\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ $[$\loneg MayAccess$^{Dir,Eve}$(o,o') \loand\linebreak
	\ablock \loforall X*\loin{Obj}, X*\loneq o. \loneg MayAccess$^{Dir,Eve}$(X*,o')$]$ \linebreak
	\ablock \loimplies\loneg MayAffect$^{Eve}$(o,o')\linebreak
\hrule
\end{logic}
With this clarification, we then reason about the relationship between protection of state (MayAffect) and paths (MayAccess$^{Ind,*}$). The question we ask is what is the relation between state protection in eventual outcomes, and present path configurations? We build this relationship from the basics by progressively finding stronger conditions of MayAffect. We begin with the first condition, which says that for an object to affect another, an \textit{eventual} path has to exist. This can be derived from the definitions of our predicates.
\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Path Condition 1 (NPC1)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o') \loimplies\linebreak
	\ablock MayAccess$^{Ind,Eve}$(o,o')
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Eve}$(o,o') \loimplies\linebreak
	\ablock \loneg MayAffect$^{Eve}$(o,o')
\linebreak
\hrule
\end{logic}
Here, we immediately see a first defensive outcome of the OCap model. Having no eventual paths from object o to o' guarantees that the state of object o' cannot be modified by object o.\\

So far our results serve as a good base to enforce \loneg MayAffect$^{Eve}$(o,o') but we require \textit{stronger} necessary conditions for MayAffect$^{Eve}$(o,o'), because we need to understand the present path configurations that can deny o from modifying o'. Luckily, we have already built the necessary ingredients in our formalizations of OCap rules in \ref{ocaprules}, where have painstakingly derive the relationship between present path configurations and eventual paths. Since \textsc{npc1} tells us that the eventual paths is a necessary condition for object state modification, and \textsc{epc1} tells us the the relationship between eventual paths and present path configurations, we combine the two to present:

\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Path Condition 2 (NPC2)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')$]$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ 
\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
\ablock \loimplies \linebreak
	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
\hrule
\end{logic}
, where \textsc{npc2} is similar to \textsc{epc2}, where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use \textsc{epc2} to expand MayAccess$^{Ind,Eve}$(X*,o') till we reach the terminating MayAccess$^{Ind,Eve}$(o',o'). The final result through such as expansion is similar to \textsc{epc3***}. We use \textsc{epc3***} to make explicit the relationship between a configuration of direct capabilities between objects at \losigma, and the ability to modify state at some arising \losigma', which we call \textsc{ncc3***}:
\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Capability Condition 3*** (NCC3***)}\\
M,\losigma\ \loturns\ MayAffect$^{Ind,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*, x$^{1...n}$\loin Obj.$\{[$MayAccess$^{Ind,Now}$(o,o)\linebreak
\ablock \ablock \ablock \ablock \ablock \loand\linebreak
\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(o,x$^1$)\linebreak
\ablock \ablock \ablock \ablock \ablock $\vdots$\linebreak
\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(x$^{n-1}$,x$^n$)\linebreak
\ablock \ablock \ablock \ablock \ablock \loand\linebreak
\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(x$^n$,X*)$]$\linebreak
\ablock \ablock \ablock \ablock \quad\: \loor \linebreak
\ablock \ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$\}$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')\linebreak\\

\ablock \textit{where if o exists in \losigma,} \linebreak
\ablock \loturns\ MayAccess$^{Ind,Now}$(o,o) is true\linebreak\linebreak\
\ablock \textit{where if o' exists in some \losigma',}\linebreak
\ablock \loturns\ MayAccess$^{Ind,Eve}$(o',o') is true\linebreak
\hrule
\end{logic}

\textsc{ncc3***} gives us the present \textit{capability} configurations that will allow an eventual path from o to o', which is a necessary condition for o to modify o', and tells us that we can enforce protection of o' from o by denying all eventual paths from o to o'. But if those objects must \textit{always} and at all states be path-isolated, how do we allow them to cooperate?

\subsection{What about Cooperation?}

It is good to know we can enforce protection through some present path configurations that can guarantee eventual path-isolation, but in practice this is still not entirely useful. The quintessential question in OCap systems is whether we can enforce protection such as controlling whether o may affect o', on \textit{some condition}. In our framework, this just means that cooperation between objects require paths to exist based on some condition. This is made possible, because of Rule 3 that says objects may have private encapsulation of state and behaviour, which enables an object to deny paths leading in and leading out of it. Can our framework incorporate this?\\

How do we then express these internal object security policies in our logical framework?

It is worth revisiting our general specifications for the MayAccess$^{Dir,Eve}$ predicate, which is extremely \textit{weak}, because they have no special conditions:
\begin{logic}[MayAccess$^{Dir,Eve}$]
M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(x,e) \loiff \linebreak
\ablock \loexists \losigma' \loin Arising(M,\losigma).\linebreak
\ablock \loexists f. \loexec{x.f}{\losigma'} = \loexec{e}{\losigma'}
\loor \linebreak
\ablock (\losigma'(this) = \loexec{x}{\losigma'} \loand\ \loexists y. \losigma'(y) = \loexec{e}{\losigma'})\linebreak
\ablock \loand\ $\{\}$\texttt{ // No conditions}
\end{logic}

The reason why we have not used any special conditions is because this also allows us to generalize any kind of object which we do not know how it behaves. The most dangerous kind of objects are those that exposes all their behaviours to be public, and can expose all capabilities that is possesses through those behaviours. In that end, our general specifications are meant to assume a worst-case scenario, and hence it is no surprise that we insist on path-isolation to guarantee protection.\\

But, what if we have code knowledge of some set of objects in our system so we know how they behave?\\
